Disentangling the spinal mechanisms of illusory heat and burning sensations in the Thermal Grill Illusion

### 2023-07-14 Manuscript.Rmd
updated results

In keeping with the unique heat and burning features of TGI, our results exhibited a more robust TGI when stimuli were confined within dermatomes compared to when applied across dermatomes, corresponding to non-adjacent spinal segments (Fig. 2).
We investigated this difference in subjective cold / warm and burning ratings with a two by two interaction of manipulation (control / TGI) and dermatome (within / across).
Participants reported a reduced experience of cold, when the stimulus was presented within determatomes compared to across in the TGI condition compared to the control condition when the reference thermode was cold (experiment 1), but not when the reference thermode was warm (experiment 2) (Exp. 1: $\beta$ = `r stats_model_cold_exp1$beta[6]`, `r make_pvalue(stats_model_cold_exp1$p[6])`; Exp. 2: $\beta$ = `r stats_model_cold_exp2$beta[6]`, `r make_pvalue(stats_model_cold_exp2$p[6])`) The percept of warm showed no siginifcant interaction with either thermode as reference (Exp. 1: $\beta$ = `r stats_model_warm_exp1$beta[6]`, `r make_pvalue(stats_model_warm_exp1$p[6])`; Exp. 2: $\beta$ = `r stats_model_warm_exp2$beta[6]`, `r make_pvalue(stats_model_warm_exp2$p[6])`) The percept of burning showed no statistical significant interaction with the reference being cold, but a strong interaction when the reference thermode was warm (Exp.1: $\beta$ = `r stats_model_burn_exp1$beta[6]`, `r make_pvalue(stats_model_burn_exp1$p[6])`; Exp. 2 $\beta$ = `r stats_model_burn_exp2$beta[6]`, `r make_pvalue(stats_model_burn_exp2$p[6])`) These findings not only corroborated earlier results regarding the segmental organisation of the TGI [@fardo_organization_2018] but also suggested that both the thermosensory enhancement and illusory pain components of the illusion are modulated at the spinal cord level.
These findings are consistent with previous results on the modulation of heat and pain ratings of TGI stimuli by conditioned pain modulation [@harper_conditioned_2017].
All together these results support the role of spinal processes in the generation of distinct perceptual aspects of the TGI.

>>>>

In keeping with the unique heat and burning features of TGI, our results exhibited a more robust TGI when stimuli were confined within dermatomes compared to when applied across dermatomes, corresponding to non-adjacent spinal segments (Fig. 2).
When rating the cold thermode (Exp. 1), participants reported a significantly reduced subjective experience of cold for TGI, but not non-TGI, stimuli applied within a dermatome compared to across dermatomes (stimulation by dermatome interaction: $\beta$ = `r stats_model_cold_exp1$beta[6]`, `r make_pvalue(stats_model_cold_exp1$p[6])`.
However we did not observed such interaction effect for warm (stimulation by dermatome interaction: $\beta$ = `r stats_model_warm_exp1$beta[6]`, `r make_pvalue(stats_model_warm_exp1$p[6])`) and burning ratings (stimulation by dermatome interaction: $\beta$ = `r stats_model_burn_exp1$beta[6]`, `r make_pvalue(stats_model_burn_exp1$p[6])`; .

In contrast, when rating the warm thermode (Exp. 2), participants reported markedly enhanced burning sensations for TGI, but not non-TGI, stimuli applied within a dermatome compared to across dermatomes (stimulation by dermatome interaction: $\beta$ = `r stats_model_burn_exp2$beta[6]`, `r make_pvalue(stats_model_burn_exp2$p[6])`).
However, we did not observe a modulation of cold (stimulation by dermatome interaction: $\beta$ = `r stats_model_cold_exp2$beta[6]`, `r make_pvalue(stats_model_cold_exp2$p[6])`) ) and warm thermosensory ratings (stimulation by dermatome interaction: $\beta$ = `r stats_model_warm_exp2$beta[6]`, `r make_pvalue(stats_model_warm_exp2$p[6])`).

These findings not only corroborated earlier results regarding the segmental organisation of the TGI [@fardo_organization_2018] but also suggested that both the thermosensory enhancement and illusory pain components of the illusion are modulated at the spinal cord level.
These findings are consistent with previous results on the modulation of heat and pain ratings of TGI stimuli by conditioned pain modulation [@harper_conditioned_2017].
All together these results support the role of spinal processes in the generation of distinct perceptual aspects of the TGI.
